---
title: "Introduction to Data Science"
subtitle: "Homework 2"
author: "Logan Zorrilla - `lzorrilla0358@floridapoly.edu`"
output: html_notebook
---



# Problem 1 

(a) (10 points) Define a function `imperial_to_metric` that takes in two arguments: a number of feet and a number of inches. The function should return the equivalent length in meters

```{r}
# Convert imperial values to metric
feet <- 50
feet_to_meters <- feet*0.3048
print("The feet input converted to meters is:")
print(feet_to_meters)

inches <- 75
inches_to_cm <- inches*2.54
print("The inches input converted to cenimeters is:")
print(inches_to_cm)

```


(b) (5 points) Create a variable `height_in_meters` by passing your height in imperial to the `imperial_to_metric` function

```{r}
# Convert my height in inches to a height in meters
height_in_inches <- 75

height_in_meters <- height_in_inches/39.37
print("My height in meters is:")
print(height_in_meters)

```
# Problem 2

`R` comes with many functions that you can use to do sophisticated tasks like random sampling. For example, you can round a number with the `round()` function, or calculate its factorial with the `factorial()` function. Using a function is pretty simple. Just write the name of the function and then the data you want the function to operate on in parentheses.

(a) (5 points) Find `round(3.1415)`, `factorial(12)`, `mean(1:17)`.
```{r}
# Output the values using the sophisticated functions.
round(3.1415)
factorial(12)
mean(1:17)

```

(b) (5 points) You can create an object named `die` that contains the numbers one through six:

```{r}
# Create a vector with numbers from 1 to 6
die <- 1:6
```

You can simulate a roll of the die with `R`’s `sample()` function. `sample()` takes two arguments: a vector named `x` and a number named `size`. 

Try `sample(x = die, size = 1)`. 

Is there any difference if you type `sample(die, size = 1)` ?

```{r}

sample(x = die, size = 1)

```


# Problem 3

Use `R` as a calculator to compute the following values:

(a) (5 points) $27 (38 -17)$

(b) (5 points) $ln(14^7)$

(c) (5 points) $\sqrt\frac{436}{12}$

```{r}

part_a <- 27*(38-17)
print(part_a)

part_b <- log(14^7)
print(part_b)

part_c <- sqrt(436/12)
print(part_c)


```



# Problem 4

Create the following vectors in `R`:

$a = (5, 10, 15, 20, ..., 160)$

$b = (87, 86, 85, ..., 56)$

```{r}

a <- (seq(5,160, by =5))
print(seq(5,160, by =5))

b <- (seq(87, 56, by =-1))
print(seq(87, 56, by =-1))

```

(a) (5 points) Multiply these vectors and call the result `d`. 

```{r}

d <- a*b
print(d)

```

(b) (15 points) Select subsets of `d` to identify the following.

- What are the 19th, 20th, and 21st elements of `d`?

- What are all of the elements of `d` which are less than 2000?

- How many elements of `d` are greater than 6000?

```{r}

d[c(19,20,21)]

elements_a <- d[-(1:16)]
print(elements_a)

elements_b <- length(elements_a)
print(elements_b)

```


# Problem 5

Start with this vector of `nouns`.  

```{r}
nouns <- c("apple", "flower", "insect", "lettuce", "knife", "dog", "cloud", "person", "cabinet", "flower" )
```

(a) (5 points) Use the `length()` function to display the number of elements in `nouns`.

(b) (5 points) Use indexing to create a new vector consisting of the first 4 elements of `nouns`.

(c) (5 points) Use indexing to create a new vector consisting of only the last 8 elements of `nouns`.

(d) (5 points) Use indexing to create a new vector of the 1st, 3rd through 6th, and 10th elements of `nouns` (the length of the resulting vector should be 6)

(e) (5 points) Create a new vector with the elements of `nouns` in reverse order.

```{r}

length(nouns)

nouns[-(1:4)]

nouns[-(1:2)]

nouns[c(1,3:6,10)]

nouns[c(10:1)]

```

# Problem 6

Consider the function `rescale01` defined as 

```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

(a) (5 points) Evaluate `rescale01(c(0, 5, 10))`
The output would be 0.0, 0.5 and 1.0

(b) (5 points) Describe what the `rescale01` function does.

(c) (5 points) What would happen if `x` contained a single missing value, and `na.rm` was `FALSE`? Show an example.





