---
title: "Introduction to Data Science"
subtitle: "Homework 3"
author: "Logan Zorrilla - `lzorrilla0358@floridapoly.edu`"
output: html_notebook
---



# Problem 1 

Load the `tidyverse` package

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

In this problem you will work with vehicle fuel economy data from the US EPA (Environmental Protection Agency), ranging from 1985-2015. 

Read the data using the code below:

```{r, message=FALSE, warning=FALSE}
vehicles <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/vehicles.csv")
```

   Variable  | Description
-------------|-----------------------------------------
`id`         | Unique EPA identifier
`make`       | Manufacturer
`model`      | Model name
`year`       | Model year
`class`      | EPA vehicle size class
`trans`      | Transmission
`drive`      | Drive train
`cyl`        | Number of cylinders
`displ`      | Engine displacement, in litres
`fuel`       | Fuel type
`hwy`        | Highway fuel economy, in mpg
`cty`        | City fuel economy, in mpg


(a) (6 points) Select the different manufacturers (makes) of the cars in this data set. Save this vector in a variable called `makes`

```{r, message=FALSE, warning=FALSE}
makes <- select(vehicles, make)
```

(b) (6 points) Can you determine how many different car manufacturers are represented by the data set?

```{r, message=FALSE, warning=FALSE}
count(distinct(makes))
```

(c) (6 points) Filter the data set for vehicles manufactured in 1998.
```{r, message=FALSE, warning=FALSE}
filter(vehicles, year == "1998")
```
(d) (8 points) Mutate the 1998 cars data frame to add a column `average` that has the average gas mileage (between city and highway MPG) for each car
```{r, message=FALSE, warning=FALSE}
cars_98 <- mutate(vehicles, average = (hwy + cty)/2)
cars_98
```
(e) (8 points) Of all vehicles, what is the vehicle ID of the vehicle with the worst highway MPG?
```{r}
vehicles %>%
  filter(hwy == min(hwy))%>%
  select(id)
```
(f) (8 points) Which 2013 Acura model has the best highway MPG? Which 2015 Honda model has the best highway MPG?
```{r}
vehicles %>%
  filter(make == "Acura", year == "2013") %>%
  filter(hwy ==max(hwy)) %>%
  select(model)
```
```{r}
vehicles %>%
  filter(make == "Honda", year == "2015") %>%
  filter(hwy ==max(hwy)) %>%
  select(model)
```



# Problem 2

Load the `nycflights13` packages. 

```{r}
library(nycflights13)
```


Use the `flights` dataframe to answer the following questions. 
```{r}
glimpse(flights)
```

(a) (6 points) Find all flights headed to TPA, and select only the origin, destination, and the departure time.
```{r}
flights%>%
  filter(dest =="TPA") %>%
  select(origin, dest, dep_time)
```
(b) (6 points) Consider flights from JFK to MCO. What was the average, minimum, and maximum _air time_
of those flights? 
```{r}
flights%>%
  filter(origin == "JFK" & dest =="MCO") %>%
  summarise(mean = mean(air_time, na.rm = T), min = min(air_time, na.rm = T), max = max(air_time, na.rm=T))
```

(c) (6 points) What was the average _departure delay_ in each month? Save this as a data frame `dep_delay_by_month`. (Hint: you may have to perform a grouping operation then summarizing your data).
```{r}
dep_delay_by_month <- flights%>%
  group_by(month) %>%
  summarise(mean_delay =mean(dep_delay, na.rm = T))

dep_delay_by_month
```

(d) (8 points) Which month had the greatest average _departure delay_?
```{r}
dep_delay_by_month %>%
  filter(mean_delay==max(mean_delay))
```
(e) (8 points) To which destinations were the average _arrival delays_ the highest?
```{r}
flights %>%
  group_by(dest) %>%
  summarise(arr_delay =mean(arr_delay, na.rm = T)) %>%
  arrange(desc(arr_delay))
```
(f) (8 points) Which airline had the smallest average _arrival delay_?
```{r}
flights %>%
  group_by(carrier) %>%
  summarise(arr_delay =mean(arr_delay, na.rm = T)) %>%
  filter(arr_delay==min(arr_delay))
```
(g) (8 points) Find all flights that were operated by United, American, or Delta.
```{r}
flights %>%
  filter(carrier=="UA"|carrier=="AA"|carrier=="DL") %>%
  select(carrier, origin, dest, arr_delay) %>%
  arrange(desc(arr_delay))
```
(h) (8 points) Find all the flights (if any) headed to Orlando, Tampa, or Miami with a scheduled departure time later than 8:00 PM.
```{r}
flights %>%
  filter(dest=="MCO"|dest=="TPA"|dest=="MIA") %>%
  filter(sched_dep_time>2000) %>%
  select(dest, sched_dep_time) %>%
  arrange(sched_dep_time)
```



